---
title: "StatsFinal"
format: html
editor: visual
---

This project uses a dataset on video game habits among participants over a one-month period. The goal is to examine how much time people spend playing video games on weekdays compared to weekends and whether there is a statistically significant difference between them.


The dataset includes the following variables:

- `participant_id`: A unique identifier for each participant (categorical)
- `weekday_playtime`: Total average playtime on weekdays in hours (numeric)
- `weekend_playtime`: Total average playtime on weekends in hours (numeric)
- `age`: Age of participant (numeric)
- `gender`: Gender of participant (categorical: Male, Female, Other)

Research Questions

1. How does video game playtime compare between weekdays and weekends?
2. Is there a correlation between age and average weekend playtime?


Histogram of Weekend Playtime

```{r}
game_data <- data.frame(weekend_playtime = c(2.5, 3, 5, 1.5, 4, 6, 3.5, 2, 7, 5.5, 4.5, 3))
hist(game_data$weekend_playtime, 
     main = "Distribution of Weekend Playtime", 
     xlab = "Hours Played", 
     col = "lightblue", 
     border = "white")
```
Boxplot of Weekday vs. Weekend Playtime
```{r}
library(dplyr)
library(tidyr)
library(ggplot2)

print(names(game_data))  

str(game_data)

game_data$weekday_playtime <- c(2, 3, 1.5, 2, 3, 4, 3, 2, 2.5, 1.8, 3, 4.2)  

long_data <- game_data %>%
  pivot_longer(cols = c(weekday_playtime, weekend_playtime), 
               names_to = "day_type", 
               values_to = "hours_played")

ggplot(long_data, aes(x = day_type, y = hours_played, fill = day_type)) +
  geom_boxplot() +
  labs(title = "Video Game Playtime: Weekdays vs. Weekends", 
       x = "Day Type", 
       y = "Hours Played") +
  scale_fill_manual(values = c("skyblue", "lightgreen")) +
  theme_minimal()
```

Summary Statistics
```{r}
library(dplyr)

data <- data.frame(weekend_playtime = c(2.5, 3, 5, 1.5, 4, 6, 3.5, 2, 7, 5.5, 3, 4),
                   weekday_playtime = c(2, 3, 1.5, 2, 3, 4, 3, 2, 2.5, 1.8, 3, 4.2))  

str(data)  

if (!"weekday_playtime" %in% names(data)) {
  data$weekday_playtime <- c(2, 3, 1.5, 2, 3, 4, 3, 2, 2.5, 1.8, 3, 4.2)  
}

if (!"weekend_playtime" %in% names(data)) {
  data$weekend_playtime <- c(2.5, 3, 5, 1.5, 4, 6, 3.5, 2, 7, 5.5, 3, 4)
}

data %>%
  summarise(
    mean_weekday = mean(weekday_playtime, na.rm = TRUE),
    sd_weekday = sd(weekday_playtime, na.rm = TRUE),
    mean_weekend = mean(weekend_playtime, na.rm = TRUE),
    sd_weekend = sd(weekend_playtime, na.rm = TRUE)
  )
```
Paired Sample T-test
We use a paired t-test since each participant has both weekday and weekend playtime data.
```{r}
t.test(data$weekday_playtime, data$weekend_playtime, paired = TRUE)
```

Correlation: Age vs. Weekend Playtime
```{r}
if (!"age" %in% names(data)) {
  data$age <- c(25, 30, 22, 28, 35, 40, 26, 29, 32, 33, 27, 31)  
}

data$age <- as.numeric(data$age)
data$weekend_playtime <- as.numeric(data$weekend_playtime)

data <- na.omit(data)

cor.test(data$age, data$weekend_playtime)
```
Conclusion:
In this project, we analyzed a dataset of participants' video game habits to explore how playtime differs between weekdays and weekends. Descriptive statistics showed that weekend playtime tends to be higher, with a wider spread. Visualizations including histograms and boxplots helped illustrate this difference clearly.
Using a paired sample t-test, I found a statistically significant difference in playtime between weekdays and weekends, suggesting that people play significantly more on weekends. Additionally, a correlation test showed a weak negative correlation between age and weekend playtime, indicating that age may have a minor influence on weekend gaming habits. These findings highlight behavioral patterns that may be useful for game developers or digital wellbeing studies.


Video Presentation:
https://drive.google.com/file/d/1BQ-cdDRGeUzA3zixgtC48zM0txZUIKCS/view?usp=drive_link